Estimating survival in left filtered data

### 2022-03-04 paper.tex
SER

The present project replicates the simulation and bias analyses
presented in Chapter 3 of Izano (2017), embeds the problem in the
non-parametric structural causal approach of Pearl (1995), and applies
the WKM and AWKM estimators to the estimation of digestive-system
cancer-free survival under exposure and no censoring rules in the UAW-GM
Cohort. The \texttt{stremr} package was used to estimate the WKM and
AWKM survival curves in both the simulation and the applied analyses
(Sofrygin, van der Laan, and Neugebauer 2021).

@@

\begin{table}[H]
\centering
\caption{True cancer-free survival time $\mu_{a}$ over 20-year follow-up and estimator averages over 250 replicates.} 
\label{tab:survival}
\begin{tabular}{llrrrrr}
  \toprule
Rule & Estimator & Scenario 1 & Scenario 2 & Scenario 3 & Scenario 4 & Scenario 5 \\ 
  \midrule
$a_0$ & Truth & 19.08 & 19.10 & 19.10 & 18.02 & 17.83 \\ 
   & WKM & 19.52 & 19.51 & 19.52 & 18.91 & 18.90 \\ 
   & AWKM & 19.01 & 19.00 & 19.01 & 17.80 & 17.71 \\ 
   \midrule
$a_1$ & Truth & 18.80 & 18.79 & 18.76 & 17.54 & 17.29 \\ 
   & WKM & 19.41 & 19.39 & 19.39 & 18.68 & 18.67 \\ 
   & AWKM & 18.76 & 18.76 & 18.74 & 17.30 & 17.22 \\ 
   \bottomrule
\end{tabular}
\end{table}
\begin{table}[H]
\centering
\caption{Difference in average cancer-free survival time over 20-year follow-up comparing rule $a_1$ always exposed to rule $a_0$ never exposed at work: true value $\psi$ and estimator averages over 250 replicates.} 
\label{tab:differences}
\begin{tabular}{lrrrrr}
  \toprule
Estimator & Scenario 1 & Scenario 2 & Scenario 3 & Scenario 4 & Scenario 5 \\ 
  \midrule
Truth & -0.28 & -0.31 & -0.34 & -0.48 & -0.54 \\ 
  WKM & -0.11 & -0.12 & -0.13 & -0.23 & -0.23 \\ 
  AWKM & -0.25 & -0.24 & -0.27 & -0.50 & -0.50 \\ 
   \bottomrule
\end{tabular}
\end{table}
\begin{table}[H]
\centering
\caption{Bias estimates of estimators for $\psi$, the difference in average cancer-free survival time over 20 years of follow-up.} 
\label{tab:bias}
\begin{tabular}{lrrrrr}
  \toprule
Estimator & Scenario 1 & Scenario 2 & Scenario 3 & Scenario 4 & Scenario 5 \\ 
  \midrule
WKM & 0.17 & 0.19 & 0.21 & 0.25 & 0.31 \\ 
  AWKM & 0.03 & 0.07 & 0.07 & -0.02 & 0.04 \\ 
   \bottomrule
\end{tabular}
\end{table}

@@

In the simulation study, we showed that under several scenarios
compatible with our hypothetical causal structure, the AWKM survival
estimator had a smaller bias than the WKM estimator. The bias was
smaller when the cumulative incidence of the outcome was low and at
later follow-up time points. Next, we estimated cancer-free survival in
a real-world context. Using data from two plants participating in the
UAW-GM Cohort study, we followed \(26\,182\) individuals starting from
hire to 35 years after hire for incidence of digestive system cancers
(colon, rectal, esophageal, or stomach). This duration of follow-up
spans a sizable proportion of a usual worker's working lifetime. As in
the simulation, the UAW-GM data were longitudinal data with baseline
covariates, time-varying covariates, and a survival outcome.

The exposures of interest were MWF of three types: straight, soluble,
and synthetic (Byers 2006; F. Mirer 2003; F. E. Mirer 2010). Straight
MWFs are hydrocarbon-based fluids that became widely-used by the 1920s.
They continue to occupy a large portion of the MWF market due to their
simple formulation. In straight MWFs, hydrocarbons of different lengths
are mixed together with other additives to attain different properties.
Straight MWFs contain polycyclic aromatic hydrocarbons, long known to be
carcinogenic (IARC 1973). Soluble oils are water-based oil emulsions
first introduced in response to rising oil prices. They now make up the
largest market share of MWFs (Childers 2006). Soluble MWFs are
vulnerable to microbial contamination, so they contain biocides and
chlorinated chemicals. Soluble MWFs have carcinogenic potential from
both their oil components and their additives, but their high lubricity
makes them the most popular fluid type. Synthetic MWFs have the best
toxicological profile, have no oil, and have a higher resistance to
microbial growth. They were introduced into the MWF market in the second
half of the 20th C., but failed to out-perform soluble MWFs in
industrial metalworking applications. Synthetic MWFs contain biocides,
corrosion inhibitors, and chlorinated compounds, some classified as
carcinogenic by the IARC (IARC 1987).

The outcome of interest was digestive system cancer incidence. There is
little past research linking digestive system cancers to MWF exposures,
but there is some evidence suggesting that straight MWFs cause digestive
system cancers (Izano et al. 2019). Cancer incidence was obtained by
linkage to Surveillance, Epidemiology, and End Results (SEER), which
recorded cancer incidence cases starting on January 1, 1973. The cohort
is comprised of individuals hired between 1938 and 1975. Cancer-free
survival to the start of the registry was a left-filtering process
possibly in the presence of the HWSE, as was the case in simulations.
Over the 35 year follow-up period, vital status was obtained through the
Social Security Administration, the National Death Index, as well as
records provided by the UAW. The exposure rules of interest in the
applied analysis were different than those in the simulation study:
\(a_0\) being exposed with 75\% chance in the years exposure was
observed and \(a_1\) being exposed in the years exposure was observed.
Weights were truncated at 1000. Exposure was binary and defined to be
whether there was exposure above the median level of exposure to
straight, soluble, and synthetic MWFs. Counterfactual survival under
rules \(a_0\) and \(a_1\) were estimated using the WKM and AWKM
estimators. Treatment and censoring mechanisms were estimated using
logistic regression conditional on baseline and time-varying
confounders. These logistic regressions were estimated with
stratification by every two years of follow-up. Baseline confounders
included race, sex, plant, and year of hire. Time-varying confounders
included age, cumulative time off, employment status, and exposure to
the metalworking fluids exposure from past years (3 years in the case of
exposure and 6 years in the case of death due to censoring). Summary
statistics for the full study population and those who experienced
digestive system cancer are presented in Table \ref{tab:popchar}.

@@

Estimated digestive system cancer-free curves under rules \(a_0\) and
\(a_1\) applied to the three MWF types are presented in Figure
\ref{fig:gm_survival}. Surprisingly, the survival curves were more or
less overlapping under the two exposure rules. Numeric values for the
cancer-free survival and difference in cancer-free survival are
presented in Tables \ref{tab:gm_survival} and \ref{tab:gm_differences}.
The numeric summaries were consistent with the qualitative
interpretation of the estimated survival curves; expected survival time
does not differ substantially comparing rules \(a_0\) and \(a_1\)
applied to all MWF types. Nonetheless, reducing exposure to straight and
soluble MWFs yielded a longer estimated survival time, though the
difference was less than a month in time.

\begin{figure}[h]
\caption{Cancer-free survival over time since hire under interventions for exposure to straight, soluble, and synthetic metalworking fluids. The estimated inverse probability weighted Kaplan-Meier (WKM) survival curve is represented by the solid. The estimated Aalen-filtered inverse probability weighted Kaplan-Meier (AWKM) survival curve is represented by the dashed-line. Salmon color indicates survival and survival estimates under rule $a_0$; cyan color indicates those under rule $a_1$.}
\label{fig:gm_survival}
\begin{center}
\includegraphics{/Users/kevinchen/Documents/2021 Fall/256 - Causal Inference/256-final/reports/resources/gm_results.pdf}
\end{center}
\end{figure}

\begin{table}[H]
\centering
\caption{Estimated cancer-free survival time over 35-year follow-up.} 
\label{tab:gm_survival}
\begin{tabular}{llrrr}
  \toprule
Rule & Estimator & Straight & Soluble & Synthetic \\ 
  \midrule
$a_0$ & WKM & 29.9606 & 29.9605 & 29.9626 \\ 
  $a_0$ & AWKM & 29.9337 & 29.9398 & 29.9443 \\ 
   \midrule
$a_1$ & WKM & 29.9600 & 29.9584 & 29.9648 \\ 
  $a_1$ & AWKM & 29.9325 & 29.9360 & 29.9479 \\ 
   \bottomrule
\end{tabular}
\end{table}
\begin{table}[H]
\centering
\caption{Estimated difference in average cancer-free survival time over 35-year follow-up comparing rule $a_1$ exposed above the median level of exposure with probability 1\% to $a_0$ never exposed.} 
\label{tab:gm_differences}
\begin{tabular}{lrrr}
  \toprule
Estimator & Straight & Soluble & Synthetic \\ 
  \midrule
WKM & -0.0006 & -0.0021 & 0.0022 \\ 
  AWKM & -0.0012 & -0.0039 & 0.0037 \\ 
   \bottomrule
\end{tabular}
\end{table}

Figure \ref{fig:gm_weights} presents the median cumulative weight
applied years since hire with ribbons showing the minimum and maximum
weight. The overlap in the distribution of weights under the two rules
for synthetic MWF exposure suggests that the model for the treatment and
censoring mechanisms were inadequate in distinguishing the units of
analysis by their probability of following a certain exposure rule or of
remaining alive. The distribution of the weights was very skewed.
Without truncation, they would have been in the order of magnitude of
\(10^{20}\) or even higher. Distributional summaries of the cumulative
weights without truncation are presented in Table \ref{tab:gm_w}. The
presence of extremely large weights suggests that our observed data were
inadequate for answering the causal questions of interest due to
practical violations in overlap (Petersen et al. 2012).

\begin{table}[H]
\centering
\caption{Summary statistics for the cumulative weights used for the WKM estimator at $t = 1$ and $t = 35$ years since hire for each metalworking fluid type and exposure rule.} 
\label{tab:gm_w}
\begin{tabular}{llrrrrrrr}
  \toprule
MWF type & Rule & $t$ & Minimum & Q1 & Median & Mean & Q3 & Maximum \\ 
  \midrule
Straight & $a_0$ &   1 & 1.00 & 1.00 & 1.00 &   1 & 1.00 &   1 \\ 
   & $a_1$ &   1 & 1.00 & 1.00 & 1.00 &   1 & 1.00 &   1 \\ 
  Soluble & $a_0$ &   1 & 1.00 & 1.00 & 1.00 &   1 & 1.00 &   1 \\ 
   & $a_1$ &   1 & 1.00 & 1.00 & 1.00 &   1 & 1.00 &   1 \\ 
  Synthetic & $a_0$ &   1 & 1.00 & 1.00 & 1.00 &   1 & 1.00 &   1 \\ 
   & $a_1$ &   1 & 1.00 & 1.00 & 1.00 &   1 & 1.00 &   1 \\ 
  Straight & $a_0$ &  35 & 0.00 & 2.25 & 4.71 & $9.7 \times 10^{16}$ & 4622.00 & $7.3 \times 10^{20}$ \\ 
   & $a_1$ &  35 & 0.00 & 2.70 & 13.40 & $1.3 \times 10^{18}$ & 22142.45 & $9.8 \times 10^{21}$ \\ 
  Soluble & $a_0$ &  35 & 0.00 & 4.04 & 858.96 & $5.2 \times 10^{15}$ & 74610.19 & $2.3 \times 10^{19}$ \\ 
   & $a_1$ &  35 & 0.00 & 158.56 & 6685.12 & $2.3 \times 10^{17}$ & 1227858.13 & $1.7 \times 10^{21}$ \\ 
  Synthetic & $a_0$ &  35 & 0.00 & 1.02 & 1.25 & $3.7 \times 10^{19}$ & 8.85 & $3.2 \times 10^{23}$ \\ 
   & $a_1$ &  35 & 0.00 & 1.02 & 1.36 & $2.1 \times 10^{21}$ & 22.63 & $1.8 \times 10^{25}$ \\ 
   \bottomrule
\end{tabular}
\end{table}

\begin{figure}
\caption{Median of the cumulative weights for each year since hire with ribbons delimiting the range of values calculated. Salmon color corresponds to the weights used in the estimation of the survival curve under rule $a_0$; cyan corresponds to those for rule $a_1$.}
\label{fig:gm_weights}
\begin{center}
\includegraphics{/Users/kevinchen/Documents/2021 Fall/256 - Causal Inference/256-final/reports/resources/gm_weights.pdf}
\end{center}
\end{figure}

@@

\leavevmode\hypertarget{ref-IARC_1973}{}%
IARC. 1973. \emph{IARC Monographs on the Evaluation of Carcinogenic Risk
of the Chemical to Man: Certain Polycyclic Aromatic Hydrocarbons and
Heterocyclic Compounds}. Vol. 3. World Health Organization International
Agency for Research on Cancer.

\leavevmode\hypertarget{ref-IARC_1987}{}%
---------. 1987. \emph{IARC Monographs on the Evaluation of the
Carcinogenic Risk of Chemicals to Humans. Overall Evaluations of
Carcinogenicity: An Updating of IARC Monographs, Volumes 1 to 42}.
\emph{World Health Organization}. Vol. 7. World Health Organization
International Agency for Research on Cancer.

>>>>

The present project replicates the simulation and bias analyses
presented in Chapter 3 of Izano (2017), embeds the problem in the
non-parametric structural causal approach of Pearl (1995), and applies
the WKM and AWKM estimators to the estimation of digestive-system
cancer-free survival under exposure and no censoring rules in the UAW-GM
Cohort. The \texttt{stremr} package was used to estimate the WKM and
AWKM survival curves in both the simulation and the applied analyses
(Sofrygin, van der Laan, and Neugebauer 2021). All code necessary for
replicating the analyses presented here may be found on
\href{https://github.com/kvntchn/gm-delayed-entry.git}{GitHub/kvntchn/gm-delayed-entry}.
Note that running the script for the applied analyses will require
additional permissions.

@@

\begin{table}[H]
\centering
\caption{True cancer-free survival time $\mu_{a}$ over 20-year follow-up and estimator averages over 250 replicates.} 
\label{tab:survival}
\begin{tabular}{llrrrrr}
  \toprule
Rule & Estimator & Scenario 1 & Scenario 2 & Scenario 3 & Scenario 4 & Scenario 5 \\ 
  \midrule
$a_0$ & Truth & 19.08 & 19.10 & 19.10 & 18.02 & 17.83 \\ 
   & WKM & 19.52 & 19.51 & 19.52 & 18.91 & 18.90 \\ 
   & AWKM & 19.01 & 19.00 & 19.01 & 17.80 & 17.71 \\ 
   \midrule
$a_1$ & Truth & 18.80 & 18.79 & 18.76 & 17.54 & 17.29 \\ 
   & WKM & 19.41 & 19.39 & 19.39 & 18.68 & 18.67 \\ 
   & AWKM & 18.76 & 18.76 & 18.74 & 17.30 & 17.22 \\ 
   \bottomrule
\end{tabular}
\end{table}
\begin{table}[H]
\centering
\caption{Difference in average cancer-free survival time over 20-year follow-up comparing rule $a_1$ always exposed to rule $a_0$ never exposed at work: true value $\psi$ and estimator averages over 250 replicates.} 
\label{tab:differences}
\begin{tabular}{lrrrrr}
  \toprule
Estimator & Scenario 1 & Scenario 2 & Scenario 3 & Scenario 4 & Scenario 5 \\ 
  \midrule
Truth & -0.28 & -0.31 & -0.34 & -0.48 & -0.54 \\ 
  WKM & -0.11 & -0.12 & -0.13 & -0.23 & -0.23 \\ 
  AWKM & -0.25 & -0.24 & -0.27 & -0.50 & -0.50 \\ 
   \bottomrule
\end{tabular}
\end{table}
\begin{table}[H]
\centering
\caption{Bias estimates of estimators for $\psi$, the difference in average cancer-free survival time over 20 years of follow-up.} 
\label{tab:bias}
\begin{tabular}{lrrrrr}
  \toprule
Estimator & Scenario 1 & Scenario 2 & Scenario 3 & Scenario 4 & Scenario 5 \\ 
  \midrule
WKM & 0.17 & 0.19 & 0.21 & 0.25 & 0.31 \\ 
  AWKM & 0.03 & 0.07 & 0.07 & -0.02 & 0.04 \\ 
   \bottomrule
\end{tabular}
\end{table}
\begin{table}[H]
\centering
\caption{True cancer-free 20-year risk and estimator averages over 250 replicates.} 
\label{tab:survival}
\begin{tabular}{llrrrrr}
  \toprule
Rule & Estimator & Scenario 1 & Scenario 2 & Scenario 3 & Scenario 4 & Scenario 5 \\ 
  \midrule
Rule $a_0$ & Truth & 0.12 & 0.12 & 0.12 & 0.25 & 0.26 \\ 
  Rule $a_0$ & WKM & 0.07 & 0.07 & 0.07 & 0.15 & 0.15 \\ 
  Rule $a_0$ & AWKM & 0.10 & 0.10 & 0.10 & 0.23 & 0.23 \\ 
   \midrule
Rule $a_1$ & Truth & 0.15 & 0.15 & 0.15 & 0.29 & 0.31 \\ 
  Rule $a_1$ & WKM & 0.08 & 0.08 & 0.08 & 0.18 & 0.18 \\ 
  Rule $a_1$ & AWKM & 0.13 & 0.13 & 0.13 & 0.27 & 0.27 \\ 
   \bottomrule
\end{tabular}
\end{table}

@@

In the simulation study, we showed that under several scenarios
compatible with our hypothetical causal structure, the AWKM survival
estimator had a smaller bias than the WKM estimator. The bias was
smaller when the cumulative incidence of the outcome was low and at
later follow-up time points. Next, we estimated cancer-free survival in
a real-world context. Using data from two plants participating in the
UAW-GM Cohort study, we followed \(26\,182\) individuals starting from
hire to 55 years after hire for incidence of digestive system cancers
(colon, rectal, esophageal, or stomach). As in the simulation, the
UAW-GM data were longitudinal data with baseline covariates,
time-varying covariates, and a survival outcome.

The exposures of interest were MWF of three types: straight, soluble,
and synthetic (Byers 2006; F. Mirer 2003; F. E. Mirer 2010). Straight
MWFs are hydrocarbon-based fluids that became widely-used by the 1920s.
They continue to occupy a large portion of the MWF market due to their
simple formulation. In straight MWFs, hydrocarbons of different lengths
are mixed together with other additives to attain different properties.
Straight MWFs contain polycyclic aromatic hydrocarbons, long known to be
carcinogenic (\emph{IARC Monographs on the Evaluation of Carcinogenic
Risk of the Chemical to Man: Certain Polycyclic Aromatic Hydrocarbons
and Heterocyclic Compounds} 1973). Soluble oils are water-based oil
emulsions first introduced in response to rising oil prices. They now
make up the largest market share of MWFs (Childers 2006). Soluble MWFs
are vulnerable to microbial contamination, so they contain biocides and
chlorinated chemicals. Soluble MWFs have carcinogenic potential from
both their oil components and their additives, but their high lubricity
makes them the most popular fluid type. Synthetic MWFs have the best
toxicological profile, have no oil, and have a higher resistance to
microbial growth. They were introduced into the MWF market in the second
half of the 20th C., but failed to out-perform soluble MWFs in
industrial metalworking applications. Synthetic MWFs contain biocides,
corrosion inhibitors, and chlorinated compounds, some classified as
carcinogenic by the IARC (IARC 1987).

The outcome of interest was digestive system cancer incidence. There is
little past research linking digestive system cancers to MWF exposures,
but there is some evidence suggesting that straight MWFs cause digestive
system cancers (Izano et al. 2019). Cancer incidence was obtained by
linkage to Surveillance, Epidemiology, and End Results (SEER), which
recorded cancer incidence cases starting on January 1, 1973. The cohort
is comprised of individuals hired between 1938 and 1985. Cancer-free
survival to the start of the registry was a left-filtering process
possibly in the presence of the HWSE, as was the case in simulations.
Over the 55-year follow-up period, vital status was obtained through the
Social Security Administration, the National Death Index, as well as
records provided by the UAW. The exposure rules of interest in the
applied analysis were different than those in the simulation study:
\(a_0\) having cumulative exposure greater than the lower threshold as
observed and \(a_1\) having no exposure. Weights were truncated at 500.
Counterfactual survival under rules \(a_0\) and \(a_1\) were estimated
using the WKM and AWKM estimators. Treatment and censoring mechanisms
were estimated using logistic regression conditional on baseline and
time-varying confounders. These logistic regressions were estimated with
stratification by every two years of follow-up. Baseline confounders
included race, sex, plant, and year of hire. Time-varying confounders
included age, cumulative time off, employment status, and cumulative
exposure to the metalworking fluids exposure up to through previous
year. Summary statistics for the full study population and those who
experienced digestive system cancer are presented in Table
\ref{tab:popchar}.

@@

Estimated digestive system cancer-free curves under rules \(a_0\) and
\(a_1\) applied to the three MWF types are presented in Figure
\ref{fig:gm_survival}. Numeric values for the cancer-free survival and
difference in cancer-free survival are presented in Tables
\ref{tab:gm_survival} and \ref{tab:gm_differences}. The numeric
summaries were consistent with the qualitative interpretation of the
estimated survival curves. Nonetheless, reducing exposure to straight
and soluble MWFs yielded a longer expected digestive-free survival time.

\begin{figure}[h]
\caption{Cancer-free survival over time since hire under interventions for exposure to straight, soluble, and synthetic metalworking fluids. The estimated inverse probability weighted Kaplan-Meier (WKM) survival curve is represented by the solid. The estimated Aalen-filtered inverse probability weighted Kaplan-Meier (AWKM) survival curve is represented by the dashed-line. Salmon color indicates survival and survival estimates under rule $a_0$; cyan color indicates those under rule $a_1$.}
\label{fig:gm_survival}
\begin{center}
\includegraphics{/Users/kevinchen/eisen/gm-delayed-entry/reports/resources/gm_results.pdf}
\end{center}
\end{figure}

\begin{table}[H]
\centering
\caption{Estimated cancer-free survival time over 55-year follow-up.} 
\label{tab:gm_survival}
\begin{tabular}{llrrr}
  \toprule
Rule & Estimator & Straight & Soluble & Synthetic \\ 
  \midrule
As observed & WKM & 53.8388 & 54.0067 & 54.1434 \\ 
  As observed & AWKM & 53.7402 & 53.9111 & 54.0855 \\ 
   \midrule
No exposure & WKM & 54.1002 & 54.2562 & 53.9552 \\ 
  No exposure & AWKM & 54.0099 & 54.1731 & 53.8377 \\ 
   \bottomrule
\end{tabular}
\end{table}
\begin{table}[H]
\centering
\caption{Estimated difference in average cancer-free survival time over 55-year follow-up comparing rule $a_1$ exposed above the median level of exposure with probability 1\% to $a_0$ never exposed.} 
\label{tab:gm_differences}
\begin{tabular}{lrrr}
  \toprule
Estimator & Straight & Soluble & Synthetic \\ 
  \midrule
WKM & 0.2615 & 0.2495 & -0.1883 \\ 
  AWKM & 0.2697 & 0.2620 & -0.2478 \\ 
   \bottomrule
\end{tabular}
\end{table}

Distributional summaries of the cumulative weights without truncation
are presented in Table \ref{tab:gm_w}. The presence of extremely large
weights suggests that our observed data were inadequate for answering
the causal questions of interest due to practical violations in overlap
(Petersen et al. 2012).

\begin{table}[H]
\centering
\caption{Summary statistics for the cumulative weights used for the WKM estimator at $t = 1$ and $t = 55$ years since hire for each metalworking fluid type and exposure rule.} 
\label{tab:gm_w}
\begin{tabular}{llrrrrrrr}
  \toprule
MWF type & Rule & $t$ & Minimum & Q1 & Median & Mean & Q3 & Maximum \\ 
  \midrule
Straight & $a_0$ &   1 &   1 &   1 &   1 &   1 &   1 &   1 \\ 
   & $a_1$ &   1 &   1 &   1 &   1 &   1 &   1 &   1 \\ 
  Soluble & $a_0$ &   1 &   1 &   1 &   1 &   1 &   1 &   1 \\ 
   & $a_1$ &   1 &   1 &   1 &   1 &   1 &   1 &   1 \\ 
  Synthetic & $a_0$ &   1 &   1 &   1 &   1 &   1 &   1 &   1 \\ 
   & $a_1$ &   1 &   1 &   1 &   1 &   1 &   1 &   1 \\ 
  Straight & $a_0$ &  55 &   0 & 5.5 & 9.9 &  51 &  36 & $2.4 \times 10^{3}$ \\ 
   & $a_1$ &  55 &   0 &   0 &   0 & 3.6 & 5.1 & $1.4 \times 10^{2}$ \\ 
  Soluble & $a_0$ &  55 &   0 & 4.2 & 5.9 &  52 &  22 & $6 \times 10^{3}$ \\ 
   & $a_1$ &  55 &   0 &   0 &   0 & 5.5 &   0 & $1.7 \times 10^{3}$ \\ 
  Synthetic & $a_0$ &  55 &   0 & 3.2 & 5.6 &  34 &  18 & $2.8 \times 10^{3}$ \\ 
   & $a_1$ &  55 &   0 &   0 & 2.9 & 3.5 & 4.3 & $1 \times 10^{2}$ \\ 
   \bottomrule
\end{tabular}
\end{table}

\begin{figure}
\caption{Median of the cumulative weights for each year since hire with ribbons delimiting the range of values calculated. Salmon color corresponds to the weights used in the estimation of the survival curve under rule $a_0$; cyan corresponds to those for rule $a_1$.}
\label{fig:gm_weights}
\begin{center}
\includegraphics{/Users/kevinchen/eisen/gm-delayed-entry/reports/resources/gm_weights.pdf}
\end{center}
\end{figure}

@@

\leavevmode\hypertarget{ref-IARC_1987}{}%
IARC. 1987. \emph{IARC Monographs on the Evaluation of the Carcinogenic
Risk of Chemicals to Humans. Overall Evaluations of Carcinogenicity: An
Updating of IARC Monographs, Volumes 1 to 42}. \emph{World Health
Organization}. Vol. 7. World Health Organization International Agency
for Research on Cancer.

\leavevmode\hypertarget{ref-IARC_1973}{}%
\emph{IARC Monographs on the Evaluation of Carcinogenic Risk of the
Chemical to Man: Certain Polycyclic Aromatic Hydrocarbons and
Heterocyclic Compounds}. 1973. Vol. 3. World Health Organization
International Agency for Research on Cancer.

### 2022-03-04 reports/03-results.Rmd
SER

```{r}
sim_results <- readRDS(here("reports/resources", "sim_results.rds"))
survival.tab <- dcast(
	sim_results[,.(
		Survival = sum(Survival)
	), .(Rule = rule, scenario, Estimator)],
	... ~ scenario,
	value.var = "Survival"
)

@@

print(xtable(
	bias.tab,
	caption = "Bias estimates of estimators for $\\psi$, the difference in average cancer-free survival time over 20 years of follow-up.",
	label = "tab:bias"),
	floating = T)
```

>>>>

```{r}
sim_results <- readRDS(here("reports/resources", "sim_results.rds"))
risk.tab <- dcast(
	sim_results[t == max(t),.(
		Risk = 1 - Survival
	), .(Rule = rule, scenario, Estimator)],
	... ~ scenario,
	value.var = "Risk"
)

setorder(risk.tab, Rule, Estimator)

survival.tab <- dcast(
	sim_results[,.(
		Survival = sum(Survival)
	), .(Rule = rule, scenario, Estimator)],
	... ~ scenario,
	value.var = "Survival"
)

@@

print(xtable(
	bias.tab,
	caption = "Bias estimates of estimators for $\\psi$, the difference in average cancer-free survival time over 20 years of follow-up.",
	label = "tab:bias"),
	floating = T)

print(xtable(risk.tab,
			 caption = "True cancer-free 20-year risk and estimator averages over 250 replicates.",
			 label = "tab:survival"),
			hline.after = c(-1,0,3, nrow(survival.tab)),
			floating = T)
```
